### 2016-08-08 Docs/managing_jobs/managingjobs.tex
document gpu profiling at OLCF

>>>>

\subsection{Profiling and Debugging on GPUs}
To get an idea of how code performs on Titan's GPUs, there are a few tools
available.  We'll overview a few here.

\subsubsection{{\tt Score-P} with {\tt CUBE} and {\tt vampir}}

{\tt Score-P} is a profiling and tracing tool that can be used to instrument
code to generate data for other tools to analyze, such as {\tt CUBE} and 
{\tt vampir}.  These tools have been developed to analyze performance of HPC
codes that run on several nodes, not specifically for analyzing GPU usage.
Still, they do support some GPU analysis.  In the next section, we'll discuss
NVIDA's tools specifically for analyzing GPU usage.
At the time of writing, {\tt Score-P} usage is fairly well documented on OLCF's
website here: \url{https://www.olcf.ornl.gov/kb_articles/software-scorep/}.
We'll review the essentials here, but please see the link for more details.

To instrument a code with {\tt Score-P} you must re-compile.  First, your
desired modules will need to be loaded.  Please note that \emph{order is
important} --- you need to load modules needed for compilation before loading 
{\tt Score-P}.  The {\tt Score-P} module is designed to detect the loaded
environment and will configure itself based on that.  These tools have been
tested with the PGI 16.3.0 compilers, and we will use them in our examples.
One possible set of module loads is

\begin{lstlisting}[language=bash,mathescape=false]
$ module load PrgEnv-pgi
$ module swap pgi/15.7.0 pgi/16.3.0
$ module load cudatoolkit
$ module load scorep/3.0-rc1
\end{lstlisting}

In the above we've loaded version 3.0, release candidate 1, which added some
support for analyzing OpenACC code.  The next step is to compile.  You
essentially preface your normal compile (and link) line with the {\tt Score-P}
executable and options.  As an example using the Fortran wrapper required on
Titan, we have

\begin{lstlisting}[language=bash,mathescape=false]
$ scorep --cuda --openacc -v ftn program.f90
\end{lstlisting}

One way to achieve this in \maestro\ is to modify the appropriate make file.  For
PGI, this would be {\tt \$BOXLIB\_HOME/Tools/F\_mk/comps/Linux\_pgi.mak}.  If
this proves useful, it may be worth it to build {\tt Score-P} into the build
infrastructure.

Once compiled, we are ready to generate profiling and tracing data.  Among those
that develop these tools, note that they draw a distinction between profiling
and tracing.  Profiling generates a timing (or perhaps other metric) summary of
the entire program's execution while tracing produces a timeline of the
execution. {\tt Score-P}'s analysis is configured with environment variables.
Some of the key configuration variables used in testing include

\begin{lstlisting}[language=bash,mathescape=false]
export SCOREP_ENABLE_PROFILING=yes
export SCOREP_ENABLE_TRACING=yes
export SCOREP_EXPERIMENT_DIRECTORY=profile-trace
export SCOREP_CUDA_ENABLE=yes,kernel_counter,flushatexit
export SCOREP_CUDA_BUFFER=200M
export SCOREP_TOTAL_MEMORY=1G
export SCOREP_OPENACC_ENABLE=yes
\end{lstlisting}

For a full listing and definition of configuration variables, execute
\begin{lstlisting}[language=bash,mathescape=false]
$ scorep-info config-vars --full 
\end{lstlisting}

Except for very simple codes, you will never want to enable both tracing and
profiling.  The overhead is too high, and the code will likely crash or be
excessively slow.  Typically, it's best to profile first and then trace.  The
profiling data can be used to help configure tracing (as we'll see shortly).

Once the configuration is set, simply run the code as you normally would.
Experience suggests you will need to load the same modules that were loaded for
compilation when executing.  If analysis is being done through a batch script,
note that you cannot do a simple {\tt module load ...} in the script.  First you
need to do {\tt source /opt/modules/default/init/bash} in the script, and then
module loads will work as usual.

After executing, analysis data will be stored in the specified
{\tt SCOREP\_EXPERIMENT\_DIRECTORY}.  With profiling, you
will see a file like profile.cubex.  This can be opened with cube 
(module load cube).

As mentioned, the profiling data can be used to get
recommended settings for tracing.  Running
\begin{lstlisting}[language=bash,mathescape=false]
$ scorep-score -r profile.cubex
\end{lstlisting}
will yield output showing estimated sizes for e.g.~{\tt SCOREP\_TOTAL\_MEMORY}.  It
can also suggest functions to that are called a lot to be filtered out of the
trace.  If you don't care about them and they're slowing scorep down (or making
an outrageously large output file), this is useful to do.

Running with tracing enabled will generate a traces.otf2 file that can be
inspected with vampir (module load vampir)

\subsubsection{{\tt nvprof} and {\tt nvvp}}

NVIDIA provides tools for specifically analyzing how your code utilizing their
GPUs.  {\tt Score-P} is a fully-featured profiling with some CUDA and OpenACC
support.  It can be useful for providing context for GPU execution and it allows
you to, for example, see line numbers for OpenACC directives that are executed.
{\tt nvprof} will only analyze GPU execution, but in exchange you get much more
detail than is available with {\tt Score-P}.  {\tt nvvp} is NVIDIA's visual
profiler.  It can be used to read data generated by {\tt nvprof}.  Most useful
is the guided analysis it will perform, which analyzes your code's GPU
performance for bottlenecks and suggests ways to improve performance.  Both are
provided when you load the {\tt cudatoolkit} module.

With {\tt nvprof}, no instrumentation is necessary.  Instead, you compile
normally and then run {\tt nvprof} on the executable.  As before, be sure when
executing to load the modules used for compilation.  Executing {\tt nvprof} on
Titan's compute nodes requires some unexpected options having to do with how
{\tt aprun} and {\tt nvprof} interact.

To get a basic overview printed to the terminal on Titan's compute node, execute
\begin{lstlisting}[language=bash,mathescape=false]
$ aprun -b nvprof --profile-child-processes ./gpuprogram.exe arg1 arg2... 
\end{lstlisting}

To generate tracing data for {\tt nvvp}, execute
\begin{lstlisting}[language=bash,mathescape=false]
$ aprun -b nvprof --profile-child-processes -o nvprof.timeline.out%p 
  ./gpuprogram.exe arg1 arg2... 
\end{lstlisting}
{\tt nvvp} can then be used to read {\tt nvprof.timeline.out\%p}, where the 
{\tt \%p} will be replaced with the process ID.

To generate profile-like metric data for {\tt nvvp}, execute
\begin{lstlisting}[language=bash,mathescape=false]
$ aprun -b nvprof --profile-child-processes --analysis-metrics 
  -o nvprof.metrics.out%p ./gpuprogram.exe arg1 arg2... 
\end{lstlisting}
This is the output needed for {\tt nvvp}'s guided analysis.

\subsubsection{Target Metrics}
The output from profilers may be difficult to makes sense of.  The purpose of
this section is to note different metrics and reasonable targets for them.
Note that these may be specific to the k20x hardware in Titan.
\begin{itemize}
   \item Threads per block: 256-512.  Note that if your code requires many
      registers per thread, then this will limit the number of threads per block.
   \item Occupancy: 60\% is a reasonable target.  We have had success with codes
      even achieving only 23\% occupancy.  
\end{itemize}
